On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-10-29 paper.tex
Cleaning
of repeated or obsolete things
Rewriting to avoid inconsistency of coordinates
In some parts of the text and figures the "actual" coordinates (ia, jb) were used, whereas in others the normalized coordinates (i,j) were used. They were even mixed in the same sentence of figure. This has been unified in the text for consistency. The actual coordinates are now used. The choice had to be this, otherwise the intuition of Euclidean distance was lost. Pending: correct labels in Figures 6 and 8 according to this

%  \ldots ***maybe a short paragraph here mentioning related problems: Buffon's needle, Buffon-Laplace problem. Some references: \cite{Ramaley69}, \cite{Arnow94}, \cite{Mathai99}. 
% Alex: Mentioned Ramaley69 and Arnow94, not sure where Mathai99 fits in
%\url{https://www.jstor.org/stable/2317945}
%\url{https://www.jstor.org/stable/2687085}

@@

Consider $a=b=1$ first. This is illustrated in Figure~\ref{fig: ijLS_1}. Note that in this and in the next figures the axes represent $ia$ and $jb$, rather than normalized coordinates $i$ and $j$. In this graph, each dashed diagonal line contains pairs $(i,j)$ with the same $i+j-1$; and the lower bound \eqref{eq: len > sqrt} corresponds to an arc centered at normalized coordinates $(2,2)$.

@@

For a given $\len \in \mathbb R^+$, the $(i,j)$ pairs that can be achieved with segments of length not exceeding $\len$ are, by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, those that satisfy \eqref{eq: len > sqrt}. Thus for a given value of $i+j-1$ the optimal $(i,j)$ pair (the one that can be achieved with the smallest possible length) is that with minimum distance to the point with normalized coordinates $(2,2)$. Such pairs are depicted as filled circles in the figure. The displayed arcs represent the inequality \eqref{eq: len > sqrt} for each of the optimal pairs. Conversely, given a length $\len$, the maximum number of visited tiles will be achieved with one of these pairs, namely that situated on the highest (or rightmost) diagonal line that still lies within the circle with radius $\len$ centered at normalized coordinates $(2,2)$.

@@

As a second example, consider $a=1.35$, $b=1$. This is depicted in Figure~\ref{fig: ijLS_1p35}. Again, the optimal pair $(i,j)$ for each diagonal is that closest to $(2,2)$; but in this case the $i,j$ coordinates of these pairs do not follow a rule as simple as in the previous example.

@@

Consider, for the moment, $i$, $j$ as if they were real-valued variables, and denote $x=ia$, $y=jb$. The line $i+j-1 = \tiles$ then becomes
\begin{equation}
\label{eq: line 1}
\frac x a + \frac y b = \tiles + 1,
\end{equation}
and $(i-2)^2 a^2 + (j-2)^2 b^2$ is expressed as $(x-2a)^2+(y-2b)^2$. The point minimizing this quadratic function along \eqref{eq: line 1} is the intersection of this line with the perpendicular line passing through $(2a,2b)$, that is,
\begin{equation}
\label{eq: line 2}
y = \frac {a (x-2a)} b + 2b.
\end{equation}
This is depicted for $a=1.35$, $b=1$, $\tiles=7$ in Figure~\ref{fig: min suff set, form}, where \eqref{eq: line 2} is shown with solid line. Solving the system of equations \eqref{eq: line 1} and \eqref{eq: line 2} gives 
\begin{align}
\frac x a - 2 &= \frac{(t-3)b^2}{a^2+b^2}, \\
\frac y b - 2 &= \frac{(t-3)a^2}{a^2+b^2}.
\end{align}
In terms of the real-valued variables $i$, $j$, the solution $(\isolr, \jsolr)$ is
\begin{align}
\label{eq: prop: min suff set, form: isolr}
\isolr &= \frac{(t-3)b^2}{a^2+b^2} + 2, \\
\label{eq: prop: min suff set, form: jsolr}
\jsolr &= \frac{(t-3)a^2}{a^2+b^2} + 2.
\end{align}
This point is shown in Figure~\ref{fig: min suff set, form} with a square marker.

% ***This figure and Figure~\ref{fig: ijLS_1p35_detail} label the points as (i,j) instead of (ia,jb). Clarify in the text, or change the figures ***Maybe use the concept of normalized coordinates
\begin{figure}%
\centering%
\includegraphics[width=.7\textwidth]{mss}%
\caption{Obtaining $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ in  Proposition~\ref{prop: min suff set, form}. Example with $a=1.35$, $b=1$, $\tiles=7$%
}%
\label{fig: min suff set, form}%
\end{figure}%

The variables $i$, $j$ are actually limited to integer values. The point $(i_\tiles, j_\tiles)$ that minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ along the line $i+j-1=\tiles$ restricted to integer values is either $(\lceil \isolr \rceil, \lfloor \jsolr \rfloor)$ or $(\lfloor \isolr \rfloor, \lceil \jsolr \rceil)$, whichever is closest to $(\isolr, \jsolr)$, as can be seen in Figure~\ref{fig: min suff set, form}. This can be expressed as
\begin{align}
\label{eq: prop: min suff set, form: isoli}
i_\tiles &= \left\lfloor \isolr + \frac 1 2 \right\rfloor, \\
\label{eq: prop: min suff set, form: jsoli}
j_\tiles &= \left\lceil \jsolr - \frac 1 2 \right\rceil,
\end{align}
which in case of tie takes the first of the two points indicated above. Combining \eqref{eq: prop: min suff set, form: isolr}--\eqref{eq: prop: min suff set, form: jsoli}
% *!*
yields \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}.

From \eqref{eq: prop: min suff set, form: isolr} and \eqref{eq: prop: min suff set, form: isoli},
\begin{equation}
\label{eq: prop: min suff set, form: isoli ineq}
\frac{(t-3)b^2}{a^2+b^2} + \frac 3 2 < i_\tiles \leq \frac{(t-3)b^2}{a^2+b^2} + \frac 5 2,
\end{equation}
and similarly, from \eqref{eq: prop: min suff set, form: jsolr} and \eqref{eq: prop: min suff set, form: jsoli},
\begin{equation}
\label{eq: prop: min suff set, form: jsoli ineq}
\frac{(t-3)a^2}{a^2+b^2} + \frac 3 2 \leq j_\tiles < \frac{(t-3)a^2}{a^2+b^2} + \frac 5 2.
\end{equation}
Considering the first inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the second in \eqref{eq: prop: min suff set, form: jsoli ineq} as equalities and eliminating $\tiles$ gives \eqref{eq: upper bound, line}. The point $(i_\tiles, j_\tiles)$ is strictly below that line because the considered inequalities are strict. Similarly, the bounding line \eqref{eq: lower bound, line} results from the second inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the first in \eqref{eq: prop: min suff set, form: jsoli ineq}, and the fact that both inequalities are not strict implies that the bound \eqref{eq: lower bound, line} can actually be attained.
\end{proof}

@@

For $a^2/b^2$ arbitrary, the number of pairs with the same $i$ coordinate, or with the same $j$, in the set $\mss$ is in general irregular, because the lines \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} do not follow a ``natural'' direction of the grid. This happens for instance in Figure~\ref{fig: pairs_bounds_1p35}, where the number of pairs for the first $i$ values equals either $2$ or $3$ without a clear pattern.\footnote{Strictly, there is a periodic pattern whenever $a^2/b^2$ is rational, which is the case in Figure~\ref{fig: pairs_bounds_1p35}. However, this is not easily discernible unless $a^2/b^2$ is a ratio of small numbers.}
On the other hand, a simple pattern arises when $a^2/b^2$ or $b^2/a^2$ is a natural number, as seen in Figures~\ref{fig: pairs_bounds_1} and \ref{fig: pairs_bounds_sqrt2}.

A segment whose discrete bounding rectangle has normalized width $i$ and height $j$ is oriented with approximate slope $jb/(ia)$ with respect to the $x$ axis (see Figure~\ref{fig: len ineq i j}); and the approximation becomes better for greater segment lengths. From \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} it can be seen that the positions of the pairs $(i,j) \in \mss$ have $j/i \approx a^2/b^2$ for large $i, j$. Therefore the optimal slope for long segments is approximately $a/b$. This formalizes the intuition that to maximize the number of visited tiles, the segment should follow a direction along which the perceived ``length'' of the tile is smaller.
% but not ``smallest'', if that were the case it would always be either vertical or horizontal

@@

\begin{proof}
The approach is similar to that used in Proposition~\ref{prop: min suff set, form}. Namely, for $i, j \geq 2$ the intersection $(\isolr, \jsolr)$ of the line \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ is found, and from that the actual integer solution $(\isoli, \jsoli)$ is computed.

More specifically, assuming that \eqref{eq: lower bound, line} holds, the largest $i$ that can be achieved with lengths not exceeding $\len$ is obtained, as given by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}. This value, $\isoli$, is computed by rounding down $\isolr$. Then, assuming that $i = \isoli$, the largest $j$ allowed by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, $\jsoli$, is obtained. As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in $\mss$, and in other cases it is not. However, in either case $(\isoli,\jsoli)$ has the largest $i+j-1$ sum that can be achieved with segments of length up to $\len$. Thus the desired result is $\isoli+\jsoli-1$.

The two possibilities are illustrated in Figure~\ref{fig: ijLS_1p35_detail} for $a=1.35$, $b=1$. In each case, the arc displayed in the graph is centered at $(2a,2b)$ and has radius $\len$. The inner region defined by the arc contains all $(i,j)$ pairs that are achievable according to Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}. As in previous figures, filled circles represent $(i,j)$ pairs that are in $\mss$, and empty circles are those that do not belong to $\mss$. The solid line is \eqref{eq: lower bound, line}. The intersection point $(\isolr, \jsolr)$ is displayed with a square marker.

\begin{figure}
\centering%
\subfigure[$\len = 3.1$: $(\isoli,\jsoli) \in \mss$]{%
\label{fig: ijLS_1p35_detail_in}%
\includegraphics[width=.7\textwidth]{ijLS_1p35_detail_in}%
}\\%\hfill%
\subfigure[$\len = 3.7$: $(\isoli,\jsoli) \notin \mss$]{%
\label{fig: ijLS_1p35_detail_notin}%
\includegraphics[width=.7\textwidth]{ijLS_1p35_detail_notin}%
}%
\caption{Obtaining $(\isoli,\jsoli)$ in Theorem~\ref{theo: funt, form}. Examples with $a=1.35$, $b=1$%
}%
\label{fig: ijLS_1p35_detail}%
\end{figure}%

The point $(\isolr, \jsolr)$ can be obtained as a solution of the equation system
\begin{align}
\label{eq: icont jcont 1}
(\isolr-2)^2 a^2 + (\jsolr-2)^2 b^2 &= \len^2, \\
\label{eq: icont jcont 2}
\jsolr = \frac{\isolr a^2}{b^2} - \frac{5a^2}{2b^2} + \frac 3 2.
\end{align}
Substituting \eqref{eq: icont jcont 2} into \eqref{eq: icont jcont 1} yields a quadratic equation for $\isolr$,
\begin{equation}
\label{eq: quadratic, i}
4a^2(\isolr-2)^2 - 4a^2(\isolr-2) + a^2+b^2-\frac{4\len^2 b^2}{a^2+b^2},
\end{equation}
which can have zero, one or two real-valued solutions (this respectively means that the line is exterior, tangent or secant to the circle \eqref{eq: icont jcont 1}). Only the solution with $\isolr, \jsolr \geq 2$, if any, is of interest. This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as
\begin{equation}
\label{eq: icont jcont 3}
\isolr = \frac{a + b \sqrt{4\len^2/(a^2+b^2)-1}}{2a} + 2.
\end{equation}

@@

The three cases are unified by taking the real part of \eqref{eq: icont jcont 3} and computing $\isoli = \lceil \isolr \rceil-1$, as is easily checked. This gives \eqref{eq: theo: funt, form; i}. Once $\isoli$ is known, $\jsoli$ is computed, according to \eqref{eq: theo: funt, form; j}, as the greatest integer such that $(\isoli,\jsoli)$ is within the arc determined by $\len$. This ensures that $(\isoli,\jsoli)$ is achievable with lengths less than $\len$.

There are two possibilities for the point $(\isoli,\jsoli)$, as stated at the outset. These are illustrated in Figures~\ref{fig: ijLS_1p35_detail_in} and \ref{fig: ijLS_1p35_detail_notin} respectively. The first possibility is that $(\isoli,\jsoli) \in \mss$ (Figure~\ref{fig: ijLS_1p35_detail_in}). Then, by construction  $(\isoli,\jsoli)$ maximizes $i+j-1$ among all achievable pairs of $\mss$, and is therefore optimal.

The second possibility is that $(\isoli,\jsoli) \notin \mss$ (Figure~\ref{fig: ijLS_1p35_detail_notin}). This happens when the point from $\mss$ that has $i = \isoli$ ($(4,5)$ in the figure) is outside the arc, i.e.~it would require a length greater than $\len$. The selected point  $(\isoli,\jsoli)$ ($(4,4)$ in the figure), however, has the same $i+j-1$ sum as the point from $\mss$ that ``should'' be used, which is $(\isoli-1,\jsoli+1)$ ($(3,5)$ in the figure); and therefore gives the same result. This is always the case, because  $(\isoli,\jsoli+1) \in \mss$ (it is above or on the bounding line) and $(\isoli,\jsoli) \notin \mss$ (it is below the line), and due to how $\mss$ has been constructed, this implies that $(\isoli-1,\jsoli+1) \in \mss$ and $(\isoli-1,\jsoli+k) \notin \mss$ for $k =2, 3, \ldots$.  Therefore, $(\isoli,\jsoli)$ is achievable and maximizes $i+j-1$, which implies that $\isoli+\jsoli-1$ is the desired solution.

Therefore, whether $(\isoli,\jsoli) \in \mss$ or $(\isoli,\jsoli) \notin \mss$, equations \eqref{eq: theo: funt, form; i} and \eqref{eq: theo: funt, form; j} give $\isoli+\jsoli-1$ equal to $\funt(\len)$, as claimed.

>>>>

@@

Consider $a=b=1$ first. This is illustrated in Figure~\ref{fig: ijLS_1}. Note that in this and in the next figures the axes represent $ia$ and $jb$, rather than
% normalized coordinates
$i$ and $j$. In this graph, each dashed diagonal line contains pairs $(ia,jb)$ with the same $i+j-1$; and the lower bound \eqref{eq: len > sqrt} corresponds to an arc centered at $(2a,2b)$.

@@

For a given $\len \in \mathbb R^+$, the $(i,j)$ pairs that can be achieved with segments of length not exceeding $\len$ are, by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, those that satisfy \eqref{eq: len > sqrt}. Thus for a given value of $i+j-1$ the optimal $(i,j)$ pair (the one that can be achieved with the smallest possible length) is determined by the condition that the point $(ia,jb)$ minimizes the distance to $(2a,2b)$. Such pairs are depicted as filled circles in the figure. The displayed arcs represent the inequality \eqref{eq: len > sqrt} for each of the optimal pairs. Conversely, given a length $\len$, the maximum number of visited tiles will be achieved with one of these pairs, namely the pair $(i,j)$ such that $(ia,jb)$ is on the highest (or rightmost) diagonal line while still being contained in the circle with radius $\len$ centered at $(2a,2b)$.
% ***rephrase this better?

@@

As a second example, consider $a=1.35$, $b=1$. This is depicted in Figure~\ref{fig: ijLS_1p35}. Again, the optimal pair $(i,j)$ for each diagonal is that for which the point $(ia,jb)$ is closest to $(2a,2b)$; but in this case the $i$, $j$ coordinates of these pairs do not follow a rule as simple as in the previous example.

@@

Consider, for the moment, $i$, $j$ as if they were real-valued variables, and denote $x=ia$, $y=jb$. The line $i+j-1 = \tiles$ then becomes
\begin{equation}
\label{eq: line 1}
\frac x a + \frac y b = \tiles + 1,
\end{equation}
and $(i-2)^2 a^2 + (j-2)^2 b^2$ is expressed as $(x-2a)^2+(y-2b)^2$. The point minimizing this quadratic function along \eqref{eq: line 1} is the intersection of that line with the perpendicular line passing through $(2a,2b)$, that is,
\begin{equation}
\label{eq: line 2}
y = \frac {a (x-2a)} b + 2b.
\end{equation}
This is depicted for $a=1.35$, $b=1$, $\tiles=7$ in Figure~\ref{fig: min suff set, form}, where \eqref{eq: line 2} is shown with solid line. Solving the system of equations \eqref{eq: line 1} and \eqref{eq: line 2} gives 
\begin{align}
\frac x a - 2 &= \frac{(t-3)b^2}{a^2+b^2}, \\
\frac y b - 2 &= \frac{(t-3)a^2}{a^2+b^2}.
\end{align}
In terms of the real-valued variables $i$, $j$, the solution $(\isolr, \jsolr)$ is
\begin{align}
\label{eq: prop: min suff set, form: isolr}
\isolr &= \frac{(t-3)b^2}{a^2+b^2} + 2, \\
\label{eq: prop: min suff set, form: jsolr}
\jsolr &= \frac{(t-3)a^2}{a^2+b^2} + 2.
\end{align}
The corresponding point $(\isolr a, \jsolr b)$ is shown in Figure~\ref{fig: min suff set, form} with a square marker.

% ***Correct the two labels in the figure
\begin{figure}%
\centering%
\includegraphics[width=.7\textwidth]{mss}%
\caption{Obtaining $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ in  Proposition~\ref{prop: min suff set, form}. Example with $a=1.35$, $b=1$, $\tiles=7$%
}%
\label{fig: min suff set, form}%
\end{figure}%

The variables $i$, $j$ are actually limited to integer values. The pair $(i_\tiles, j_\tiles)$ that minimizes $(i-2)^2 a^2 + (j-2)^2 b^2$ along the line $i+j-1=\tiles$ restricted to integer values is either $(\lceil \isolr \rceil, \lfloor \jsolr \rfloor)$ or $(\lfloor \isolr \rfloor, \lceil \jsolr \rceil)$, whichever gives $(i_\tiles a, j_\tiles b)$ closest to $(\isolr a, \jsolr b)$, as illustrated in Figure~\ref{fig: min suff set, form}. This can be expressed as
\begin{align}
\label{eq: prop: min suff set, form: isoli}
i_\tiles &= \left\lfloor \isolr + \frac 1 2 \right\rfloor, \\
\label{eq: prop: min suff set, form: jsoli}
j_\tiles &= \left\lceil \jsolr - \frac 1 2 \right\rceil,
\end{align}
which in case of tie takes the first of the two pairs indicated above. Combining \eqref{eq: prop: min suff set, form: isolr}--\eqref{eq: prop: min suff set, form: jsoli}
% *!*
yields \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}.

From \eqref{eq: prop: min suff set, form: isolr} and \eqref{eq: prop: min suff set, form: isoli},
\begin{equation}
\label{eq: prop: min suff set, form: isoli ineq}
\frac{(t-3)b^2}{a^2+b^2} + \frac 3 2 < i_\tiles \leq \frac{(t-3)b^2}{a^2+b^2} + \frac 5 2,
\end{equation}
and similarly, from \eqref{eq: prop: min suff set, form: jsolr} and \eqref{eq: prop: min suff set, form: jsoli},
\begin{equation}
\label{eq: prop: min suff set, form: jsoli ineq}
\frac{(t-3)a^2}{a^2+b^2} + \frac 3 2 \leq j_\tiles < \frac{(t-3)a^2}{a^2+b^2} + \frac 5 2.
\end{equation}
Considering the first inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the second in \eqref{eq: prop: min suff set, form: jsoli ineq} as equalities and eliminating $\tiles$ gives \eqref{eq: upper bound, line}. The pair $(i_\tiles, j_\tiles)$ is strictly below the line \eqref{eq: upper bound, line} because the used inequalities are strict. Similarly, the bounding line \eqref{eq: lower bound, line} results from the second inequality in \eqref{eq: prop: min suff set, form: isoli ineq} and the first in \eqref{eq: prop: min suff set, form: jsoli ineq}, and the fact that both inequalities are not strict implies that the bound \eqref{eq: lower bound, line} can actually be attained.
\end{proof}

@@

For $a^2/b^2$ arbitrary, the number of pairs in $\mss$ with the same $i$ coordinate, or with the same $j$, is in general irregular, because the lines \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} do not follow a ``natural'' direction of the grid. This happens for instance in Figure~\ref{fig: pairs_bounds_1p35}, where the number of pairs for each $i$ equals either $2$ or $3$ without a clear pattern.\footnote{Strictly, there is a periodic pattern whenever $a^2/b^2$ is rational, which is the case in Figure~\ref{fig: pairs_bounds_1p35}. However, it is not easily discernible unless $a^2/b^2$ is a ratio of small numbers.}
On the other hand, a simple pattern arises when $a^2/b^2$ or $b^2/a^2$ is a natural number, as seen in Figures~\ref{fig: pairs_bounds_1} and \ref{fig: pairs_bounds_sqrt2}.

A segment whose discrete bounding rectangle has normalized width $i$ and height $j$ is oriented with approximate slope $jb/(ia)$ with respect to the $x$ axis (see Figure~\ref{fig: len ineq i j}); and this approximation becomes better for greater segment lengths. From \eqref{eq: upper bound, line} and \eqref{eq: lower bound, line} it can be seen that the positions of the pairs $(i,j) \in \mss$ have $j/i \approx a^2/b^2$ for large $i, j$. Therefore the optimal slope for long segments is approximately $a/b$. This formalizes the intuition that to maximize the number of visited tiles, the segment should follow a direction along which the perceived ``length'' of the tile is smaller.
% but not ``smallest'', if that were the case it would always be either vertical or horizontal

@@

\begin{proof}
The approach is similar to that used in Proposition~\ref{prop: min suff set, form}. Namely, for $i, j \geq 2$ the intersection $(\isolr a, \jsolr b)$ of the line defined by \eqref{eq: lower bound, line} and the arc centered at $(2a,2b)$ with radius $\len$ is found, and from that the actual integer solution $(\isoli, \jsoli)$ is computed.

More specifically, assuming that \eqref{eq: lower bound, line} holds, the largest $i$ that can be achieved with lengths not exceeding $\len$ is obtained, as given by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}. This value, $\isoli$, is computed by rounding down $\isolr$. Then, assuming that $i = \isoli$, the largest $j$ allowed by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, $\jsoli$, is obtained. As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in $\mss$ as defined by Proposition~\ref{prop: min suff set, form}, and in other cases it is not. However, in either case $(\isoli,\jsoli)$ has the largest $i+j-1$ sum that can be achieved with segments of length up to $\len$. Thus the desired result is $\isoli+\jsoli-1$.

The two possibilities are illustrated in Figure~\ref{fig: ijLS_1p35_detail} for $a=1.35$, $b=1$. In each case, the arc displayed in the graph is centered at $(2a,2b)$ and has radius $\len$. The inner region defined by the arc contains all $(ia,jb)$ points such that the pair $(i,j)$ is achievable according to Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}. As in previous figures, filled circles represent points $(ia,jb)$ such that $(i,j) \in \mss$. The solid line is the bound \eqref{eq: lower bound, line}. The intersection point $(\isolr a, \jsolr b)$ is displayed with a square marker.

% ***Correct two labels in each of the two graphs
\begin{figure}
\centering%
\subfigure[$\len = 3.1$: $(\isoli,\jsoli) \in \mss$]{%
\label{fig: ijLS_1p35_detail_in}%
\includegraphics[width=.7\textwidth]{ijLS_1p35_detail_in}%
}\\%\hfill%
\subfigure[$\len = 3.7$: $(\isoli,\jsoli) \notin \mss$]{%
\label{fig: ijLS_1p35_detail_notin}%
\includegraphics[width=.7\textwidth]{ijLS_1p35_detail_notin}%
}%
\caption{Obtaining $(\isoli,\jsoli)$ in Theorem~\ref{theo: funt, form}. Examples with $a=1.35$, $b=1$%
}%
\label{fig: ijLS_1p35_detail}%
\end{figure}%

The pair $(\isolr, \jsolr)$ can be obtained as a solution of the equation system
\begin{align}
\label{eq: icont jcont 1}
(\isolr-2)^2 a^2 + (\jsolr-2)^2 b^2 &= \len^2, \\
\label{eq: icont jcont 2}
\jsolr = \frac{\isolr a^2}{b^2} - \frac{5a^2}{2b^2} + \frac 3 2.
\end{align}
Substituting \eqref{eq: icont jcont 2} into \eqref{eq: icont jcont 1} yields a quadratic equation for $\isolr$,
\begin{equation}
\label{eq: quadratic, i}
4a^2(\isolr-2)^2 - 4a^2(\isolr-2) + a^2+b^2-\frac{4\len^2 b^2}{a^2+b^2},
\end{equation}
which can have zero, one or two real-valued solutions (this respectively means that the line \eqref{eq: icont jcont 2} is exterior, tangent or secant to the circle \eqref{eq: icont jcont 1}). Only the solution with $\isolr, \jsolr \geq 2$, if any, is of interest. This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as
\begin{equation}
\label{eq: icont jcont 3}
\isolr = \frac{a + b \sqrt{4\len^2/(a^2+b^2)-1}}{2a} + 2.
\end{equation}

@@

The three cases are unified by taking the real part of \eqref{eq: icont jcont 3} and computing $\isoli = \lceil \isolr \rceil-1$, as is easily checked. This gives \eqref{eq: theo: funt, form; i}. Once $\isoli$ is known, $\jsoli$ is computed, according to \eqref{eq: theo: funt, form; j}, as the greatest integer such that $(\isoli a,\jsoli b)$ is within the circle determined by $\len$. This ensures that $(\isoli,\jsoli)$ is achievable with lengths less than $\len$.

There are two possibilities for the pair $(\isoli,\jsoli)$, as stated at the outset. These are illustrated in Figures~\ref{fig: ijLS_1p35_detail_in} and \ref{fig: ijLS_1p35_detail_notin} respectively. The first possibility is that $(\isoli,\jsoli) \in \mss$ (Figure~\ref{fig: ijLS_1p35_detail_in}). Then, by construction  $(\isoli,\jsoli)$ maximizes $i+j-1$ among all achievable pairs of $\mss$, and is therefore optimal.

The second possibility is that $(\isoli,\jsoli) \notin \mss$ (Figure~\ref{fig: ijLS_1p35_detail_notin}). This happens when the pair from $\mss$ that has $i = \isoli$ ($(4,5)$ in the figure) is outside the circle, i.e.~it would require a length greater than $\len$. The selected $(\isoli,\jsoli)$ ($(4,4)$ in the figure), however, has the same $i+j-1$ sum as the pair from $\mss$ that ``should'' be used, which is $(\isoli-1,\jsoli+1)$ ($(3,5)$ in the figure); and therefore gives the same result. This is always the case, because  $(\isoli,\jsoli+1) \in \mss$ (it is above or on the bounding line) and $(\isoli,\jsoli) \notin \mss$ (it is below the line), and due to how $\mss$ has been constructed, this implies that $(\isoli-1,\jsoli+1) \in \mss$ and $(\isoli-1,\jsoli+k) \notin \mss$ for $k =2, 3, \ldots$.  Therefore, $(\isoli,\jsoli)$ is achievable and maximizes $i+j-1$, which implies that $\isoli+\jsoli-1$ is the desired solution.

Therefore, regardless of whether $(\isoli,\jsoli)$ is in $\mss$ or not, \eqref{eq: theo: funt, form; i} and \eqref{eq: theo: funt, form; j} give $\isoli+\jsoli-1$ equal to $\funt(\len)$, as claimed.
